Group 1: Predicting the Outcomes of Tennis Tournaments

### 2019-12-12 Group1Prez.Rmd
update
Finalizing presentation

```{r setup, include=FALSE, warning=FALSE}
source("funs.R")
library(knitr)
options(kableExtra.latex.load_packages = FALSE) 
library(kableExtra)
library(tidyverse)
set.seed(400)
```

@@

- Project based off a paper by Paul K. Newton and Joseph B. Keller called Probability of Winning at Tennis 1. Theory and Data

@@

- It is noted that points in tennis are not iid for most purposes
  - This is not a bad assumption as the divergence from iid is small

@@

- Set: A set consists of 6 games and is won by the player who reaches 6 games first

@@

- Matches: Best of 3 set (for women) or 5 sets (for men)

@@

$p_A^S(0,0) = 1$, $p_A^S(i,j) = 0$ if i<0 or j<0.

@@

$$\text{P(A Winning the Tie Breaker | A Served Initially)} $$
$$=p_A^T = \sum_{j=0}^5p_A^T(7,j) +  p_A^T(6,6)\sum_{n=1}^{\infty} p_A^T(n+2, n) $$ \pause

@@

## Data being used 
Data used in paper:

@@

## 2002 U.S Open: Women
```{r, echo=FALSE}
Tourney1_Women = data.frame(`P(Winning a Rally)` = TDWomen[,1],
                            `Paper P(Winning the Tournement)` = c(.2872, .3599, .0482, .3074))
#`Our P(Winning the Tournement )` =round( pTC(TDWomen[1,1],TDWomen[2,1],TDWomen[3,1],TDWomen[4,1], 2), 4)

rownames(Tourney1_Women)<- c("S. Williams","L. Davenport",  "A. Mauresmo", "V. Williams")

T1<- table(base::sample(rownames(Tourney1_Women), 2000, prob = Tourney1_Women$Paper.P.Winning.the.Tournement., replace = TRUE))

@@

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") + 
  ylab("P(Winning the Tournement") +
  ggtitle("Which Player Won the Tournement?")

@@

## 2002 U.S Open: Men
```{r, echo=FALSE, }
Tourney1_Men = data.frame(`P(Winning a Rally)` = TDMen[,1],
                            `Paper P(Winning the Tournement)` = c(.6747, .0851, .0945, .1457))
#`Our P(Winning the Tournement )` =round( pTC(TDMen[1,1],TDMen[2,1],TDMen[3,1],TDMen[4,1], 2), 4)

rownames(Tourney1_Men)<- c("P. Sampras","S. Schalken",  "A. Agassi", "L. Hewitt")

T1<- table(base::sample(rownames(Tourney1_Men), 2000, prob = Tourney1_Men$Paper.P.Winning.the.Tournement., replace = TRUE))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") +
  ylab("P(Win the Tournement)") +
  ggtitle("Which Player Won the Tournement?")

@@

##2002 U.S Open: Men

@@

@@

WimWomen <- read_csv("2019 Womens Wimbledon.csv")

WimWomen$`P(Winning a Rally)` <- WimWomen$`Points Won on Serve`/WimWomen$`Total Points Served`

WimWomen <- as.data.frame(WimWomen)

WimWomen$`P(Winning Tournement)`<- pTC(WimWomen[1,4][1], WimWomen[2,4][1],WimWomen[3,4][1],WimWomen[3,4][1],2)

T1 <- table(base::sample(WimWomen$Name, 2000, prob = WimWomen$`P(Winning a Rally)`, replace = TRUE))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") +
  ylab("P(Win the Tournament)") +
  ggtitle("Which Player Won the Tournament?")

@@

WimMen <- read_csv("2019 Men's Wimbledon.csv")

WimMen <- as.data.frame(WimMen)

WimMen$`P(Winning a Rally)` <- WimMen$`Points Won on Serve`/ WimMen$`Total Points on Serve`

WimMen$`P(Winning Tournement)`<- pTC(WimMen[1,4][1], WimMen[2,4][1],WimMen[3,4][1],WimMen[3,4][1],2)

T1<- table(base::sample(WimMen$Name, 2000, prob = WimMen$`P(Winning a Rally)`, replace = TRUE))

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq/2000)) +
  xlab("Players") +
   ylab("P(Win the Tournement)")+
  ggtitle("Which Player Won the Tournement?")

## 2019 Wimbledon: Men 
```{r,echo=FALSE, fig.align='center'}
include_graphics("Novak.png")
```

>>>>

```{r setup, include=FALSE, warning=FALSE}
source("funs.R")
library(knitr)
opts_chunk$set(echo=FALSE)
options(kableExtra.latex.load_packages = FALSE) 
library(kableExtra)
library(tidyverse)
set.seed(400)
```

@@

- Based on a paper by Paul K. Newton and Joseph B. Keller called *Probability of Winning at Tennis 1. Theory and Data*

@@

- It is noted that points in tennis are not iid for most purposes. This is not a bad assumption as the divergence from iid is small

@@

- Set: A set is won by the player who wins 6 games first

@@

- Matches: Best of 3 sets (for women) or 5 sets (for men)

@@

$p_A^S(0,0) = 1$, $p_A^S(i,j) = 0$ if $i<0$ or $j<0$.

@@

$$\text{P(A Winning the Tie Breaker | A Served Initially)} $$
$$=p_A^T = \sum_{j=0}^5p_A^T(7,j) +  p_A^T(6,6)p_A^Rq_B^R[1 - p_A^Rp_B^R - q_A^Rq_B^R]^{-1}$$ \pause

@@

## Data
Data used in paper:

@@

## 2002 U.S Open: Women
```{r, echo=FALSE, message=F, warning=F}
W02 <- read_csv('2002 US Open tournament- Women.csv') %>%
  filter(!is.na(Name)) %>%
  mutate(pRally = NA)

for(i in 1:4) {
  W02$pRally[i] <- pr(as.numeric(W02[i,2]), as.numeric(W02[i,3]))
}

pRally <- as.numeric(W02$pRally)

W02$pTC <- as.numeric(pTC(pRally[1], pRally[2], pRally[3], pRally[4], 2)) #c(.2872, .3599, .0482, .3074))
#`Our P(Winning the Tournement )` =round( pTC(TDWomen[1,1],TDWomen[2,1],TDWomen[3,1],TDWomen[4,1], 2), 4)

T1<- table(base::sample(W02$Name, 2000, prob = W02$pTC, replace = TRUE))

@@

ggplot(data.frame(T1)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab("Players") + 
  ylab("Number of Times Won") +
  ggtitle("Which Player Won the Tournament?",
          subtitle = "2002 Women's US Open")

@@

## 2002 U.S Open: Men
```{r, echo=FALSE, message=FALSE, warning=FALSE}
mOpen <- read_csv('2002 US Open- Men.csv') %>%
  filter(!is.na(Name)) %>%
  mutate(pRally = NA)

for(i in 1:4) {
  mOpen$pRally[i] <- pr(as.numeric(mOpen[i,2]), as.numeric(mOpen[i,3]))
}

pRally <- as.numeric(mOpen$pRally)

mOpen$pTC <- as.numeric(pTC(pRally[1], pRally[2], pRally[3], pRally[4], 3))
#`Our P(Winning the Tournement )` =round( pTC(TDMen[1,1],TDMen[2,1],TDMen[3,1],TDMen[4,1], 2), 4)

T2<- table(base::sample(mOpen$Name, 2000, prob = mOpen$pTC, replace = TRUE))

Winner2 <- which(T2 == max(T2))

ggplot(data.frame(T2)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab("Players") +
  ylab("Number of Times Won") +
  ggtitle("Which Player Won the Tournament?",
          subtitle = "2002 Men's US Open")

@@

## 2002 U.S Open: Men

@@

## 2002 Wimbledon: Women

```{r, message=F, warning=F}
wWim02 <- read_csv('2002 Wimbledon -Women.csv') %>%
  filter(!is.na(Name)) %>%
  mutate(pRally = NA)

for(i in 1:4) {
  wWim02$pRally[i] <- pr(as.numeric(wWim02[i,2]), as.numeric(wWim02[i,3]))
}

pRally <- as.numeric(wWim02$pRally)

wWim02$pTC <- as.numeric(pTC(pRally[1], pRally[2], pRally[3], pRally[4], 2))

T3 <- table(base::sample(wWim02$Name, 2000, prob = wWim02$pTC, replace = TRUE))

ggplot(data.frame(T3)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab('Players') +
  ylab('Number of Times Won') +
  ggtitle('Which Player Won the Tournament?',
          subtitle = "2002 Women's Wimbledon")
```

## 2002 Wimbledon: Women

```{r, fig.align='center'}
include_graphics('Serena2.png')
```

## 2002 Wimbledon: Men

```{r, message=F, warning=F}
mWim02 <- read_csv('2002 Wimbledon- Men.csv') %>%
  filter(!is.na(Name)) %>%
  mutate(pRally = NA)

for(i in 1:4) {
  mWim02$pRally[i] <- pr(as.numeric(mWim02[i,2]), as.numeric(mWim02[i,3]))
}

pRally <- as.numeric(mWim02$pRally)

mWim02$pTC <- as.numeric(pTC(pRally[1], pRally[2], pRally[3], pRally[4], 2))

T4 <- table(base::sample(mWim02$Name, 2000, prob = mWim02$pTC, replace = TRUE))

ggplot(data.frame(T4)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab('Players') +
  ylab('Number of Times Won') +
  ggtitle('Which Player Won the Tournament?',
          subtitle = "2002 Men's Wimbledon")
```

## 2002 Wimbledon: Men

```{r, fig.align='center'}
include_graphics('Hewitt.png')
```

@@

wWim19 <- read_csv('2019 Womens Wimbledon.csv') %>%
  filter(!is.na(Name)) %>%
  mutate(pRally = NA)

for(i in 1:4) {
  wWim19$pRally[i] <- pr(as.numeric(wWim19[i,2]), as.numeric(wWim19[i,3]))
}

pRally <- as.numeric(wWim19$pRally)

wWim19$pTC <- as.numeric(pTC(pRally[1], pRally[2], pRally[3], pRally[4], 2))

T4 <- table(base::sample(wWim19$Name, 2000, prob = wWim19$pTC, replace = TRUE))

ggplot(data.frame(T4)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab('Players') +
  ylab('Number of Times Won') +
  ggtitle('Which Player Won the Tournament?',
          subtitle = "2019 Women's Wimbledon")
```

@@

mWim19 <- read_csv("2019 Men's Wimbledon.csv") %>%
  filter(!is.na(Name)) %>%
  mutate(pRally = NA)

for(i in 1:4) {
  mWim19$pRally[i] <- pr(as.numeric(mWim19[i,2]), as.numeric(mWim19[i,3]))
}

pRally <- as.numeric(mWim19$pRally)

mWim19$pTC <- as.numeric(pTC(pRally[1], pRally[2], pRally[3], pRally[4], 2))

T5 <- table(base::sample(mWim19$Name, 2000, prob = mWim19$pTC, replace = TRUE))

ggplot(data.frame(T5)) +
  geom_col(aes(x=Var1, y=Freq)) +
  xlab('Players') +
  ylab('Number of Times Won') +
  ggtitle('Which Player Won the Tournament?',
          subtitle = "2002 Men's Wimbledon")
```

```{r,echo=FALSE, fig.align='center'}
include_graphics("Novak.png")
```
